---
title: "ICD Code Checker"
author: "Seth Russell"
date: "9/26/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages
Here are the packages used in checking ICD codes - install if not already installed.

```{r packages}
#install.packages("iotools")
library(iotools)
#install.packages("stringi")
library(stringi)
#install.packages("dplyr")
library(dplyr)

```

## Note about pre-processing

The Check ICD Codes section below expects a text file containing a list of ICD codes, one per line.

Here's the general process I took to get the file with the codes in the desired format:

1. Copy and paste the section of dx codes from the .sas file (or .R or .do).
1. Remove all extra logic other than list of codes
1. Remove all single and double quotes
1. Replace commas with '\\n' (newline)
1. Remove any extra spaces before or after an ICD code


## Check ICD Codes

The following code snippet takes in the pre-processed ICD code list and outputs duplicates and codes that might be a superset of another code.

```{r process}
icds <- input.file("../pccc_validation/icd10_r_nov21.txt")

# prev <- ''
# ret <- lapply(icds, function(i) {
#   result <- stri_startswith_fixed(i, prev)
#   prev <<- i
#   result
# })
# 
# m <- data.frame(icds, as.matrix(ret))
# colnames(m) <- c('icd', 'prev_match')
# filter(m, prev_match == TRUE)


#alternate method to find duplicates or overlapping codes
unlist(lapply(icds, function(i) {
  tmp <- icds[icds != i]
  output <- tmp[grepl(paste0("^", i, ".*"), tmp)]
  # add the matched element into the output
  if(length(output) != 0)
    output <- c(i, output)
  output
}))


```

